---
title: "VDPO"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{VDPO}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



This is the VDPO (Variable Domain and Partially Observed) package vignette, where we
delve into the sophisticated realm of functional regression analysis. When doing
functional regression, the treatment of data characterized by variable domains or
partial observability demands specialized methodologies beyond conventional techniques.
It is within this context that VDPO emerges as a toolset within the R programming
environment, thanks to the work of the doctoral thesis of Pavel Hernandez Amaro, author
of the package.

This vignette serves as a structured exploration of practical applications of functional
regression analysis, specifically tailored to accommodate the nuances inherent in data
with variable domains or partial observability. We will use the main functions of the
package using data created with simulations. These data can be found in the package, if
anyone is interested in exploring them. The way to simulate this data can be found in
......

## First steps

First, we need to load the package

```{r setup}
library(VDPO)
```

As we said earlier, for this vignette we are going to use simulated date that
is already inside the package. 

We are going to start with a variable domain problem, so we are going to use the
`VDPO_example_vd` dataset, where *vd* stands for *variable domain*. Let's take a
quick look to the data:

```{r}
names(VDPO_example_vd)
```


```{r}
formula <- y ~ ffvd(X_se, nbasis = c(10, 10, 10))
res <- VDPO(formula = formula, data = data)
```

We can use the `summary` function to see a summary of the model

```{r}
summary(res)
```


With the model already fited, we can plot it using the `plot` function:

```{r}
plot(res)
```
