---
title: "Multivariate functional principal component analysis for variable domain data"
output:
  rmarkdown::html_vignette:
    toc: yes
    fig_width: 6
vignette: >
  %\VignetteIndexEntry{Multivariate functional principal component analysis for variable domain data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  error = FALSE,
  fig.cap = "",
  fig.path = "includes/VDPO-04-"
)
```

## Introduction

The `mfpca_vd` function performs a multivariate functional principal component
analysis for data observed on subject-specific (variable) domains. Each variable
is first decomposed through a variable domain FPCA, and the univariate scores are
combined into a domain-varying multivariate decomposition.

## Data generation

```{r setup}
library(VDPO)
```

We simulate two functional variables observed on variable domains. For each
subject, the matrix of observations carries `NA` beyond its domain, and a second
matrix holds the actual observation times.

```{r}
set.seed(1)
N <- 60
maxcols <- 28

make_variable <- function(seed_shift) {
  set.seed(100 + seed_shift)
  D  <- matrix(NA_real_, N, maxcols)
  Ti <- matrix(NA_real_, N, maxcols)
  for (i in seq_len(N)) {
    ni <- sample(12:maxcols, 1)
    tt <- sort(runif(ni, 0, ni))
    D[i, seq_len(ni)]  <- rnorm(1) * sin(tt / 3) + rnorm(1) * cos(tt / 5) + rnorm(ni, 0, 0.1)
    Ti[i, seq_len(ni)] <- tt
  }
  list(Data = D, Times = Ti)
}

v1 <- make_variable(1)
v2 <- make_variable(2)
```

## Estimation

The data are passed as a list of matrices. The observation times are given
through `Times` (one matrix per variable); alternatively, for equidistant
designs, a vector of domain lengths can be passed through `M_grid`.

```{r}
res <- mfpca_vd(
  Data  = list(v1$Data, v2$Data),
  Times = list(v1$Times, v2$Times),
  m_npcs = 3, u_npcs = 4, k_m = 8
)
str(res, max.level = 1)
```

## Eigenfunctions

The estimated eigenfunctions vary with the domain length. They can be displayed
at a few fixed domains as superimposed lines.

```{r, fig.width = 9, fig.height = 4}
plot(res, type = "eigenfunctions", variable = 1, components = 1:2)
```

The full picture across all domains is shown as a heatmap (time on the
horizontal axis, domain length on the vertical axis).

```{r}
plot(res, type = "heatmap", variable = 1, components = 1)
```

## Scores

The multivariate scores can be displayed colored by domain length.

```{r}
plot(res, type = "scores", components = 1:2)
```
