---
title: "Model fitting for variable domain functional data"
output:
  rmarkdown::html_vignette:
    toc: yes
    fig_width: 6
vignette: >
  %\VignetteIndexEntry{Model fitting for variable domain functional data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE, 
  warning = FALSE, 
  error = FALSE,
  fig.cap = "",
  fig.path = "includes/VDPO-02-"
)
```

## Introduction

The VDPO package provides, among other tools, methods for analyzing variable domain functional data. This vignette demonstrates how to fit variable domain functional regression models using the `vd_fit` function, which is designed to handle various types of functional and non-functional covariates in a flexible framework.

## Data Generation

```{r setup}
library(VDPO)
```

We'll start by generating sample data using the `data_generator_vd` function. This function creates simulated data with variable domain functional covariates and additional non-functional covariates if specified.

```{r}
# Generate data with functional and non-functional covariates
data <- data_generator_vd(beta_index = 1, use_x = TRUE, use_f = TRUE)
```

## Model Fitting

The `vd_fit` function is the main tool for fitting variable domain functional regression models. It supports various model specifications through a formula interface.

### Basic Model with Single Functional Covariate

Let's start with a basic model using only the functional covariate:

```{r}
data <- data_generator_vd(beta_index = 1, use_x = FALSE, use_f = FALSE)
formula <- y ~ ffvd(X_se, nbasis = c(10, 10, 10))
res <- vd_fit(formula = formula, data = data)
```

### Model with Multiple Functional Covariates

If your data contains multiple functional covariates, you can include them in the model:

```{r}
data <- data_generator_vd(
  beta_index = 1,
  use_x = FALSE,
  use_f = FALSE,
  multivariate = TRUE
)
formula <- y ~ ffvd(X_se, nbasis = c(10, 10, 10)) + ffvd(Y_se, nbasis = c(10, 20, 10))
res_multi <- vd_fit(formula = formula, data = data)
```

### Model with Functional and Non-Functional Covariates

The `vd_fit` function also supports including non-functional covariates, both linear and smooth terms:

```{r}
data <- data_generator_vd(beta_index = 1, use_x = TRUE, use_f = TRUE)
formula <- y ~ ffvd(X_se, nbasis = c(10, 10, 10)) + f(x2, nseg = 30, pord = 2, degree = 3) + x1
res_complex <- vd_fit(formula = formula, data = data)
```

In this model:

-   `ffvd(X_se, nbasis = c(10, 10, 10))` specifies the functional covariate
-   `f(x2, nseg = 30, pord = 2, degree = 3)` adds a smooth effect for `x2`
-   `x1` is included as a linear term

## Model Summary

You can obtain a summary of the fitted model using the `summary` function:

```{r}
summary(res_complex)
```

## Working with Non-Aligned Data

The `vd_fit` function can handle both aligned and non-aligned functional data. Here's an example with non-aligned data:

```{r}
data_not_aligned <- data_generator_vd(aligned = FALSE, beta_index = 1)
formula <- y ~ ffvd(X_se, nbasis = c(10, 10, 10))
res_not_aligned <- vd_fit(formula = formula, data = data_not_aligned)
```

## Additional functionality

If you need to include an offset in your model, you can use the `offset` argument:

```{r}
offset <- rnorm(nrow(data$X_se))
res_with_offset <- vd_fit(formula = formula, data = data, offset = offset)
```

### Ploting the betas

A heatmap for a specific beta of the model can be obtained by using the plot function:

```{r}
plot(res)
```

## Final remarks

The `vd_fit` function in the VDPO package provides a flexible and powerful tool for fitting variable domain functional regression models. It supports a wide range of model specifications, including multiple functional covariates, non-functional covariates, and various distribution families. By leveraging the formula interface, users can easily specify complex models tailored to their specific analysis needs.
