---
title: "Model fitting for partially observed functional data"
output:
  rmarkdown::html_vignette:
    toc: yes
    fig_width: 6
vignette: >
  %\VignetteIndexEntry{Model fitting for partially observed functional data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  error = FALSE,
  fig.cap = "",
  fig.path = "includes/VDPO-03-"
)
```

## Introduction

The VDPO package also provides methods for partially observed functional data,
where each curve is recorded only on a subset of its domain. This vignette shows
how to fit a partially observed functional regression model with the `po_fit`
function and how to inspect the estimated functional coefficient.

## Data generation

```{r setup}
library(VDPO)
```

The `data_generator_po_1d` function simulates a scalar response together with a
partially observed functional covariate.

```{r}
set.seed(123)
sim <- data_generator_po_1d(n = 150, grid_points = 80)
```

The returned list contains `noisy_curves_miss` (the observed curves, with `NA` on
the unobserved part of each domain), `missing_points` (the indices of the
unobserved points), `grid` (the common grid), `response` and `beta` (the true
functional coefficient).

## Model fitting

The functional covariate enters the model through the `ffpo` constructor, and the
model is fitted with `po_fit`.

```{r}
fit <- po_fit(
  response ~ ffpo(X = sim$noisy_curves_miss, missing_points = sim$missing_points,
                  grid = sim$grid, nbasis = c(30, 30)),
  data = list(response = sim$response, X = sim$noisy_curves_miss,
              grid = sim$grid, missing_points = sim$missing_points)
)
```

The fitted object reports the estimated functional coefficient together with the
covariance of its basis coefficients, from which pointwise confidence intervals
are obtained.

```{r}
str(fit, max.level = 1)
```

## Estimated coefficient

The element `Beta` holds, for each functional term, a data frame with the grid
(`t`), the estimated coefficient (`beta`), its standard error (`se`) and the
pointwise confidence limits (`lower`, `upper`).

```{r}
b <- fit$Beta[[1]]
head(b)
```

```{r}
matplot(b$t, cbind(b$beta, b$lower, b$upper), type = "l",
        lty = c(1, 2, 2), col = 1, xlab = "t", ylab = expression(beta(t)))
```

The bidimensional case (partially observed surfaces) follows the same interface
through `data_generator_po_2d`, `ffpo_2d` and `po_2d_fit`.
